---
title: "README"
author: "Colin G. Wilson"
date: "1/15/2019"
output:
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Dependencies
library(tidyverse)
library(rPref)
library(magrittr)
library(ggrepel)
```

## Pareto Passers

### Pareto-efficient performances through NFL history.

The repository contains scripts used to scrape from Pro-Football-Reference.com's play index finder and find the Pareto fronts for various criteria in player game logs, using Patrick Roocks's rPref package: https://www.p-roocks.de/rpref/

This project was inspired by an article written by Neil Paine and Andrea Jones-Rooy of FiveThirtyEight, which lays out the concept of Pareto efficiency and applies it to NBA stat lines: https://fivethirtyeight.com/features/explaining-james-hardens-monster-game-with-a-century-old-economic-theory/

### What is 'Pareto efficiency'?
The [Wikipedia article](https://en.wikipedia.org/wiki/Pareto_efficiency) explains the concept and its applications well. 

I've put together this graphic to explain 

```{r}
# Load data from input folder
load("input/qb_lines.Rdata")
```

